---
title: "Monte Carlo Simulation"
author: "Sukainah "
format: html
---


(Adapted from Monte Carlo Simulation: An Introduction for Engineers and Scientists by Alan Stevens, Chapter 11)

Polymers are large molecules comprised of a large number of repeated units (monomers) linked end to end. 
There are naturally occurring polymers, such as cellulose, and artificial polymers, such as polythene (polyethylene). 
A simple model of a polymer assumes its linked units are each oriented completely randomly with respect to each other. 
This results in a nearly infinite number of configurations.

Suppose the goal is to estimate the straight-line distance between the start and end of the polymer chain. 
The theoretical probability density function for this distance is $$P(R) = 4\pi R^2\left(\frac{3}{2\pi\langle r^2\rangle}\right)^{3/2}\exp\left(-\frac{3R^2}{2\langle r^2\rangle}\right), $$
where $R$ is the end-to-end distance, and the mean square position of the units is $\langle r^2\rangle = N u^2$ where $N$ is the number of units in the chain and $u$ is ithe length of each unit. 
For simplicity, we'll work with $u=1$. 


# Unconstrained Polymer Chains

The position of each link of the chain relative to the previous link can be described with angles $\theta\in[0,2\pi]$ and $\phi\in[0,pi]$ (think of $\theta$ as describing the direction in the XY plane and $\phi$ as describing the direction in the plane uniquely defined by the points $(0,0,0)$, $(x, y, 0)$, and $(x, y, z)$). In this case, $\phi$ need only cover $\pi$ radians. 

Without loss of generality, we can assume that the chain starts at the origin. 
Then, the end point in the chain can be described by a cumulative sum of a series of x, y, and z coordinates that are a function of $\theta_{n\times 1}$ and $\phi_{n\times 1}$:

$$\begin{array}{rcl} x &=& \cos(\theta)\sin(\phi)\\
y &=& \sin(\theta)\sin(\phi)\\
z &=& \cos(\phi)\end{array} $$

The distance between the origin and the end of the chain $(x,y,z)$ can be calculated as $\sqrt{x^2 + y^2 + z^2}$. 


## Single Unconstrained Link Function

Write a function that takes $n$ and returns the distance between the start and end points of a chain. 
Use vector-based operations to increase the efficiency of your function.
```{r}

polymerchain <- function(n) {

  
  # theta and phi angles
  theta <- runif(n, 0, 2 * pi)  
  phi <- runif(n, 0, pi)        
  
  # 3d component for each link
  the_x <- cos(theta) * sin(phi)
  the_y <- sin(theta) * sin(phi)
  the_z <- cos(phi)
  
  # sum x,y and z components togather
  x <- sum(the_x)
  y <- sum(the_y)
  z <- sum(the_z)
  
  # Calc end to end distance R
  R <- sqrt(x^2 + y^2 + z^2)
  
  return(R)
}








```


## Distribution of Unconstrained Links

Simulate 10000 chains and save the distribution of the end-to-end distance.

```{r}


# 
polymerchain <- function(n) {
  # vector starting (0,0,0)
  all_x <- 0
  all_y <- 0
  all_z <- 0
  
  # link togather
  for(i in 1:n) {
    # angles :theta+phi
    theta <- runif(1, 0, 2 * pi)
    phi <- runif(1, 0, pi)
    
    # calc link's direction
    the_x <- cos(theta) * sin(phi)
  the_y <- sin(theta) * sin(phi)
  the_z <- cos(phi)
    
    # Add to total position
    all_x <- all_x + the_x
    all_y <- all_y + the_y
    all_z <- all_z + the_z
  }
  
  # calc final distanation
  distance <- sqrt(all_x^2 + all_y^2 + all_z^2)
  return(distance)
}

#  simulation
set.seed(523582)  

# Simulate 10000 chains
results <- numeric(10000)  

for(i in 1:10000) {
  results[i] <- polymerchain(100)
}



```


# Constrained Polymer Chains
For real polymers, adjacent links don’t have the complete freedom of orientation allowed in the case of the ideal model. 
We can simulate a simple constrained orientation model by restricting the relative orientation of adjacent links.
We’ll prevent adjacent links from orienting themselves such that there is an acute angle smaller than, say 45° between them. 
We can do this by checking the angle between adjacent links after choosing a random orientation and repeatedly choosing the random orientation until that angle is larger than 45°.


## Single Constrained Link Function

Write a function that takes $n$ and returns the distance between the start and end points of a chain. 
Can you think of a way to use vectorized functions to accomplish this task? Why or why not? 

```{r}

```

## Distribution of Constrained Links

Simulate 10000 chains and save the distribution of the end-to-end distance.
Plot the density of the unconstrained links and the density of constrained links on the same plot -- how much has the distribution changed?


